Selecting Computations: Theory and Applications

### 2012-06-26 paper/vago.uai/paper.tex
as send by nick
with our changes

\title{Selecting Computations: Theory and Applications}

%\author{}
\author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
Computer Science Division \\  
University of California\\ 
Berkeley, CA 94720 \\ 
\And 
{\bf Solomon Eyal Shimony} and {\bf David Tolpin}  \\ 
Department of Computer Science \\ 
Ben-Gurion University of the Negev\\
Beer Sheva, Israel
}

@@

We thus propose the following hybrid sampling scheme: 
	at the \emph{root node}, sample based on the VOI estimate;
	at \emph{non-root nodes}, sample using UCT.

@@

One way to account for this reuse would be to incorporate the
`future' value of information into a VOI estimate. However, this 
approach requires a nontrivial extension of the theory of metareasoning for search.
Alternately, one can behave myopically with respect to the search tree depth:
\begin{enumerate}
\item Estimate VOI as though the information is discarded after each step,
\item Stop early if the VOI is below a certain threshold
   (see \secref{sec:control-stopping-criterion}), and
\item Save the unused sample budget for search in future states, such that
   if the nominal budget is $N$, and the unused budget in the last state
   is $N_u$, the search budget in the next state will be $N+N_u$.
\end{enumerate}
In this approach, the cost of a sample in the current state is the
VOI of increasing the budget of a future state by one sample.  It is
unclear whether this cost can be accurately estimated, but supposing
a fixed value for a given problem type and algorithm implementation
would work. Indeed, the empirical evaluation (\secref{sec:emp-go})
confirms that stopping and sample redistribution based on a learned
fixed cost  substantially improve the performance of the VOI-based
sampling policy in game tree search.

@@

The hybrid policies were compared on the game Go, a search domain
in which UCT-based MCTS has been particularly successful
\citep{Gelly.mogo}. A modified version of Pachi \citep{Braudis.pachi}, a state of the art
Go program, was used for the experiments:
\begin{itemize}
\item The UCT engine of Pachi was extended with VOI-aware sampling
  policies at the first step. 
\item The stopping criterion for the VOI-aware policy was
  modified and based solely on the sample cost, specified as
  a constant parameter. The heuristic stopping criterion for the
  original UCT policy was left unchanged.
\item The time-allocation model based on the fixed number of samples
  was modified for \textit{both the original UCT policy and the VOI-aware
  policies} such that 
  \begin{itemize}
    \item The same number of samples is available to
      the agent at each step, independently of the number of pre-simulated
      games;  
    \item If samples were unused at the current step,
      they become available at the next step. 
  \end{itemize}
\end{itemize}
While the UCT engine is not the most powerful engine of Pachi, it is still a strong
player. On the other hand, additional features of more advanced
engines would obstruct the MCTS phenomena which are the subject of
the experiment.
\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{uctvoi.pdf}
\caption{Winning rate of the VOI-aware policy in Go as a function of the cost $c$, for varying numbers of samples per ply.}
\label{fig:uctvoi}
\end{figure}
The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
15000 samples (game simulations) per ply, each experiment repeated
1000 times. \figref{fig:uctvoi} depicts a calibration experiment,
showing the winning rate of the VOI-aware policy against UCT as a function of
the stopping threshold $c$ (if the maximum VOI of a sample is below
the threshold, the simulation is stopped, and a move is chosen). Each
curve in the figure corresponds to a certain number of samples per
ply.  For the stopping threshold of $10^{-6}$, the VOI-aware policy
is almost always better than UCT, and reaches the winning rate of
64\% for 10000 samples per ply.

@@

>>>>

\title{Selecting Computations: Theory and Applications} 
\author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
Computer Science Division \\  
University of California\\ 
Berkeley, CA 94720 \\ 
\And 
{\bf David Tolpin} and {\bf Solomon Eyal Shimony} \\ 
Department of Computer Science \\ 
Ben-Gurion University of the Negev\\
Beer Sheva 84105, Israel
}

@@

We thus propose the following hybrid sampling scheme \citep{TolpinShimony.mcts}: 
	at the \emph{root node}, sample based on the VOI estimate;
	at \emph{non-root nodes}, sample using UCT.

@@

One way to account for this reuse would be to incorporate the
`future' value of information into a VOI estimate. However, this 
approach requires a nontrivial extension of the theory of metareasoning for search.
Alternately, one can behave myopically with respect to the search tree depth:
\begin{enumerate}
\item Estimate VOI as though the information is discarded after each step,
\item Stop early if the VOI is below a certain threshold
   (see \secref{sec:control-stopping-criterion}), and
\item Save the unused sample budget for search in future states, such that
   if the nominal budget is $N$, and the unused budget in the last state
   is $N_u$, the search budget in the next state will be $N+N_u$.
\end{enumerate}
In this approach, the cost $c$ of a sample in the current state is the
VOI of increasing the budget of a future state by one sample.  It is
unclear whether this cost can be accurately estimated, but supposing
a fixed value for a given problem type and algorithm implementation
would work. Indeed, the empirical evaluation (\secref{sec:emp-go})
confirms that stopping and sample redistribution based on a learned
fixed cost  substantially improve the performance of the VOI-based
sampling policy in game tree search.

@@

The hybrid policies were compared on the game Go, a search domain
in which UCT-based MCTS has been particularly successful
\citep{Gelly.mogo}. A modified version of Pachi \citep{Braudis.pachi}, a state of the art
Go program, was used for the experiments:
\begin{itemize}
\item The UCT engine of Pachi was extended with VOI-aware sampling
  policies at the first step. 
\item The stopping criterion for the VOI-aware policy was
  modified and based solely on the sample cost, specified as
  a constant parameter. The heuristic stopping criterion for the
  original UCT policy was left unchanged.
\item The time-allocation model based on the fixed number of samples
  was modified for \textit{both the original UCT policy and the VOI-aware
  policies} such that 
  \begin{itemize}
    \item Initially, the same number of samples is available to
      the agent at each step, independently of the number of pre-simulated
      games;  
    \item If samples were unused at the current step,
      they become available at the next step. 
  \end{itemize}
\end{itemize}
While the UCT engine is not the most powerful engine of Pachi, it is still a strong
player. On the other hand, additional features of more advanced
engines would obstruct the MCTS phenomena which are the subject of
the experiment.
\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{uctvoi.pdf}
\caption{Winning rate of the VOI-aware policy in Go as a function of the cost $c$, for varying numbers of samples per ply.}
\label{fig:uctvoi}
\end{figure}
The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
15000 samples (game simulations) per ply, each experiment repeated
1000 times. \figref{fig:uctvoi} depicts a calibration experiment,
showing the winning rate of the VOI-aware policy against UCT as a function of
the stopping threshold $c$ (if the maximum VOI of a sample is below
the threshold, the simulation is stopped, and a move is chosen). Each
curve in the figure corresponds to a certain number of samples per
ply.  For the stopping threshold of $10^{-6}$, the VOI-aware policy
is almost always better than UCT, and reaches the winning rate of
64\% for 10000 samples per ply.

@@

\input{acknowledgements}
